PEER ASSESSMENT 1
-----------------------------------------------------------------

Have all results show in the final output:  
  
```{r setoptions}
library(knitr)
opts_chunk$set(echo=TRUE)
```  

### A) READ IN DATA ###

Unzip and load data in R:

```{r dataload}

unzip("activity.zip")
rawdata <- read.csv("activity.csv", header = TRUE) ## Assumes that data is already in working directory

head(rawdata)

```

### B) MEAN AND MEDIAN ###

Calculate total number of steps taken each day, plot histogram, and determine the mean and median:

```{r sumsteps}

library(dplyr)

## Add up the total number of steps taken each day
sum_steps <- rawdata %>% group_by(date) %>% summarise(total_steps=sum(steps))
head(sum_steps)

```

``` {r histogram_rawdata}

## Plot histogram
hist(sum_steps$total_steps, col = "green",
     main = "Histogram of Total Number of Steps Taken Each Day",
     xlab = "Total Number of Steps Taken in a Day",
     ylab = "Frequency of Days") 

rug(sum_steps$total_steps)

```

```{r MeanMedian_calcs}

## Calculate mean & median of steps taken each day
mean_steps <- format(round(mean(sum_steps$total_steps, na.rm=TRUE),0),big.mark= ",")
print(mean_steps)

median_steps <- format(round(median(sum_steps$total_steps, na.rm=TRUE),0), big.mark = ",")
print(median_steps)

```

**The mean number of steps taken each day is `r mean_steps`;  
the median number of steps taken is `r median_steps`.**


### C) AVERAGE DAILY ACTIVITY PATTERN ###

Remove NAs from raw data:

```{r clean_rawdata}

rawdata_l <- complete.cases(rawdata)
data <- rawdata[rawdata_l,]

```

Plot a time series graph of the average steps taken in each 5-minute interval, and determine the 5-minute interval with the maximum average number of steps:

``` {r Daily_calcs}

# Avg steps by 5-minute interval
avgsteps_int <- rawdata %>% group_by(interval) %>% 
        summarise(avgsteps=mean(steps, na.rm = TRUE))
head(avgsteps_int)

# Avg steps across all 5-minute intervals
mean_5minsteps <- mean(rawdata$steps, na.rm=TRUE)
print(mean_5minsteps)

```

``` {r timeseries}

# Plot time series graph
with(avgsteps_int, plot(interval, avgsteps, type = "l", col = "blue",
        main = "Time series of average steps taken per 5-minute interval",
        xlab = "5-minute interval", ylab = "Average number of steps taken"))

abline(h = mean_5minsteps, lty = 2, col = "red")

legend("topright", lty = 1:2, col = c("blue", "red"),
       legend = c("Per 5-min interval", "Across all intervals"))

# Determine the 5-minute interval with the maximum average number of steps
row_max_avgsteps <- which.max(avgsteps_int$avgsteps)
int_max_avgsteps <- as.numeric(avgsteps_int[row_max_avgsteps,1])
print(int_max_avgsteps)

```

**The 5-minute interval with the maximum average number of steps is `r int_max_avgsteps`.**

### D) IMPUTING MISSING VALUES ###

``` {r Imputing_part1}

# Number of missing values
missing_int <- table(!is.na(rawdata$interval))
print(missing_int)
missing_date <- table(!is.na(rawdata$date))
print(missing_int)
missing_steps <- table(!is.na(rawdata$steps))
missing_steps <- format(missing_steps, big.mark=",")
print(missing_steps)

```

**Based on this, there were `r missing_steps["FALSE"]` missing values in the data.**

**For these missing values, the average of the other values within that 5-minute interval was used in place of the missing values. The new data with the missing data filled in is now in the dataset 'imputedata':**

``` {r Impute_parts2and3}

## Impute data using avg of each 5-minute interval to replace NAs
imputedata <- left_join(rawdata, avgsteps_int, by = "interval")

imputedata$steps[is.na(imputedata$steps)] <- 
        imputedata$avgsteps[is.na(imputedata$steps)]  ## Any missing value in 'steps' variable is to be replaced by the 'avgsteps' variable

imputedata <- select(imputedata, c(1:3))
head(imputedata)

```

Calculate total number of steps taken each day, plot histogram, and determine the mean and median, in similar vein to Question 1:

``` {r sumpsteps2}

## Add up the total number of steps taken each day, as per Q1
sum_steps2 <- imputedata %>% group_by(date) %>% summarise(total_steps2=sum(steps))
head(sum_steps2)

```

``` {r histogram_imputedata}
## Create a histogram of the number of steps taken each day, similarly to Q1
hist(sum_steps2$total_steps2, col = "green",
     main = "Histogram of Total Number of Steps Taken Each Day, 
     using Imputed Data",
     xlab = "Total Number of Steps Taken in a Day (using Imputed Data)",
     ylab = "Frequency of Days") 
rug(sum_steps2$total_steps2)

```

``` {r MeanMedian_imputedata}

## Calculate mean & median of steps taken each day, using imputed data
mean_steps2 <- format(round(mean(sum_steps2$total_steps2, na.rm=TRUE),0),big.mark= ",")
print(mean_steps2)

median_steps2 <- format(round(median(sum_steps2$total_steps2, na.rm=TRUE),0), big.mark = ",")
print(median_steps2)

```

**The mean number of steps taken each day is `r mean_steps2`;  
the median number of steps taken is `r median_steps2`.**

**Using this strategy of filling in the missing values, the mean and median remain relatively unchanged from the first part of the assignment.**


## E) WEEKDAY vs WEEKEND ##

Add weekday/weekend variable:

``` {r week_variable}

## Add weekday/weekend variable
imputedata$date <- as.POSIXlt(imputedata$date)
imputedata <- mutate(imputedata, day_of_week = weekdays(imputedata$date))
weekend <- c("Saturday","Sunday")

week <- ifelse(imputedata$day_of_week %in% weekend, "weekend", "weekday")
imputedata <- mutate(imputedata, week = week)
imputedata$week <- as.factor(imputedata$week)
imputedata$date <- as.factor(as.character(imputedata$date))
head(imputedata)

# Group by weekday/weekend
avgsteps_week <- imputedata %>% group_by(interval,week) %>% 
        summarise(avgsteps=mean(steps, na.rm = TRUE))
head(avgsteps_week)

```

Create graph:

```{r weekend_weekday}
library(lattice)
par(mfrow = c(1,1))

xyplot(avgsteps ~ interval | week, data = avgsteps_week, layout = c(1, 2),
       type = "l", xlab = "Interval", ylab = "Number of steps")

```
